---
title: "HW 3"
author: "Malvika Venkataraman"
date: "10/17/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# Problem 1

```{r}
#load data
library(p8105.datasets)
data("instacart")

#View(instacart)
```

## Description of Data

```{r}
#understanding the dataset
str(instacart)
nrow(instacart)
ncol(instacart)

#subsetting to the row of data to examine the variables
order_1 <- instacart[1,]
```

The dataset `instacart` records online grocery orders from Instacart users.  It contains `r nrow(instacart)` observations from `r length(unique(instacart$user_id))` unique users. The dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns. Some key variables are as follows:

* order_id: order identifier
* product_id: product identifier
* reordered: coded as 1 if this product has been ordered by the user in the past, 0 if otherwise
* product_name: name of the product
* order_number: the order sequence number for a user (1 = first, n = nth)
* days_since_prior_order: days since the last order, but capped at 30 days (NA of order_number = 1)

Let's take a look at a single row of data. In order `r order_1$order_id`, user `r order_1$user_id` ordered `r order_1$product_name` (product id = `r order_1$product_id`). The product is from the `r order_1$department` department, and was in the `r order_1$aisle` aisle. It was the `r order_1$add_to_cart_order`st product they added to their cart. This user has placed an order `r order_1$order_number` times before, and this product has been previously ordered by them (reordered = `r order_1$reordered`). The order was placed on a Thursday (order_dow = `r order_1$order_dow`), in the `r order_1$order_hour_of_day`th hour of the day (10am), and `r order_1$days_since_prior_order` days since the user last placed an Instacart order.  

### Aisles

```{r}
aisle_number = instacart %>% 
  distinct(aisle) %>% 
  nrow()
```
There are `r aisle_number` aisles.

```{r}
most_items_aisle = instacart %>% 
  count(aisle) %>% 
  arrange(desc(n)) %>%
  top_n(3)
```

The `r most_items_aisle$aisle[1]` aisle, is where the most items have been ordered from. The `r most_items_aisle$aisle[2]` aisle, is where the second most items have been ordered from. The `r most_items_aisle$aisle[3]` aisle, is where the third most items have been ordered from. 

### Aisles Plot

```{r}
aisle_plot = instacart %>%
  count(aisle) %>%
  mutate(aisle = fct_reorder(aisle, n)) %>%
  filter(n > 10000) %>%
  ggplot(aes(x = aisle, y = n)) + geom_point() + labs(title = "Number of Items Ordered in Each Aisle", x = "Aisle", y = "Number of Items") + theme(axis.text.x = element_text(angle = 90, size = rel(1), hjust = 1))

aisle_plot
```

### Aisles Table

```{r}
three_most_popular = instacart %>%
  filter(aisle == 'baking ingredients' | aisle == 'dog food care' | aisle == 'packaged vegetables fruits') %>%
  group_by(aisle, product_name) %>%
  summarize(
    count = n()) %>%
  top_n(3, count) %>%
  arrange(aisle, desc(count), product_name)
knitr::kable(three_most_popular)
```

### Pink Lady Apples and Coffee Ice Cream Table

```{r}
mean_hour_order = instacart %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  group_by(product_name, order_dow) %>%
  summarize(
    mean_hour = mean(order_hour_of_day)
  ) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
knitr::kable(mean_hour_order)
```

# Problem 2

## Load Data
```{r}
data("brfss_smart2010")

#View(brfss_smart2010)
```

## Data Cleaning
```{r}
brfss_data = brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  mutate(response = factor(response, 
                           levels = c("Poor", "Fair", "Good", "Very Good", "Excellent")))

#View(brfss_data)
```

## 7 or more loactions
```{r}
states_2002 = brfss_data %>%
  filter(year == 2002) %>%
  group_by(locationabbr) %>%
  distinct(locationdesc) %>%
  summarize(num_loc = n()) %>%
  filter(num_loc > 6) %>%
  pull(locationabbr)

states_2010 = brfss_data %>%
  filter(year == 2010) %>%
  group_by(locationabbr) %>%
  distinct(locationdesc) %>%
  summarize(num_loc = n()) %>%
  filter(num_loc > 6) %>%
  pull(locationabbr)

```

The states: `r states_2002` were observed at 7 or more locations in 2002. The states: `r states_2010` were observed at 7 or more locations in 2010.

## Excellent Dataset & Plot

```{r}
#construct the dataset
excellent_data = brfss_data %>%
  filter(response == "Excellent") %>%
  select(year, locationabbr, data_value) %>%
  group_by(year, locationabbr) %>%
  summarise(mean_value = mean(data_value, na.rm = T))
```

```{r}
#construct the spaghetti plot
excellent_plot = excellent_data %>%
  ggplot(aes(x = year, y = mean_value, color = locationabbr)) + geom_line(alpha = .5, ) + labs(title = "Average Value Over Time Within a State", y = "Average Value Within a State", color = "States")

excellent_plot
```

## Two Panel Plot

```{r}
#make dataset
two_panel_data = brfss_data %>%
  drop_na(response) %>%
  filter(locationabbr == "NY") %>%
  filter(year == 2006 | year == 2010) %>%
  select(year, locationabbr, locationdesc, data_value, response) %>%
  group_by(response) 

#make plot
my_title = "Distribution of Data Values for Responses Among Locations in NY State in 2006 and 2010"
two_panel_plot = two_panel_data %>%
  ggplot(aes(x = data_value, fill = response)) +
  geom_density() +
  facet_grid(~year) +
  labs(title = str_wrap(my_title, 60), x = "Data Value")

two_panel_plot
```

# Problem 3

## Load, Tidy and Wrangle Data
```{r}
#accel_data = read_csv("./data/accel_data.csv")
#View(accel_data)

accel_data = read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_count"
  ) %>%
  mutate(
    weekday = ifelse(day == "Saturday" | day == "Sunday", "weekend", "weekday"),
    minute = as.numeric(minute)
  )
```

## Description

In the accelometer dataset, there are `r nrow(accel_data)` rows and `r ncol(accel_data)` columns. Some variables that exist are the week number (week), the day id (day_id), the day of the week (day), the minute of the day (minute), the activity levels recorded in each minute of the day (activity_count), and a variable that distiguishes weekdays from weekends (weekday).

## Table of Total Activity Per Day

```{r}
accel_data %>%
  mutate(day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  group_by(week, day) %>%
  summarize(
    total_activity = sum(activity_count)
  ) %>%
  pivot_wider(
    names_from = day,
    values_from = total_activity
    ) %>%
  knitr::kable()
```

### Trends?

To investigate trends let's group activity level by day of the week
```{r}
level_by_day = accel_data %>%
  group_by(day) %>%
  summarize(total_activity = sum(activity_count))

knitr::kable(level_by_day)
```

We can also group activity level by week
```{r}
level_by_week = accel_data %>%
  group_by(week) %>%
  summarize(total_activity = sum(activity_count))

knitr::kable(level_by_week)
```

From our investigation, we can see that activity levels are highest on Fridays and Wednesdays, and lowest on Saturdays and Tuesdays. In our second table, we also see that activity levels were the highest in Week 2 and lowest in Week 4.


## Plot 24 Hr Activity for Each Day

```{r}
accel_data %>%
  mutate(day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>% 
  ggplot(aes(x = minute, y = activity_count, color = day)) + 
  geom_line() +
  labs(
    title = "Activity Over the Course of the Day",
    x = "Minute",
    y = "Activity"
  )
```

### Analysis
Several observations can be made from this graph. On all days of the week, activity in lower during the first ~300 minutes of the day, as well as the last ~150 minutes of the day. A possible explanation for this is that the patient may be sleeping during these times.

On certain days, the patient has different spikes in activity. On Friday, the patient's activity is higher in the night and it runs longer into the night than on other days. On Sundays, the patient's peak activity occurs around minute 850 approximately. On Thursdays, there is a spike in activity around minute 350 approximately.
