---
title: "HW 3"
author: "Malvika Venkataraman"
date: "10/17/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# Problem 1

```{r}
#load data
library(p8105.datasets)
data("instacart")

#View(instacart)
```

## Description of Data

```{r}
#understanding the dataset
str(instacart)
nrow(instacart)
ncol(instacart)
length(unique(instacart$user_id))

order_1 <- instacart[1,]

order_1$order_id
order_1$product_id
order_1$add_to_cart_order
order_1$product_name
```

The dataset `instacart` records online grocery orders from Instacart users.  It contains `r nrow(instacart)` observations from `r length(unique(instacart$user_id))` unique users. The dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns. Some key variables are as follows:

* order_id: order identifier
* product_id: product identifier
* reordered: coded as 1 if this product has been ordered by the user in the past, 0 if otherwise
* product_name: name of the product
* order_number: the order sequence number for a user (1 = first, n = nth)
* days_since_prior_order: days since the last order, but capped at 30 days (NA of order_number = 1)

Let's take a look at a single row of data. In order `r order_1$order_id`, user `r order_1$user_id` ordered `r order_1$product_name` (product id = `r order_1$product_id`). The product is from the `r order_1$department` department, and was in the `r order_1$aisle` aisle. It was the `r order_1$add_to_cart_order`st product they added to their cart. This user has placed an order `r order_1$order_number` times before, and this product has been previously ordered by them (reordered = `r order_1$reordered`). The order was placed on a Thursday (order_dow = `r order_1$order_dow`), in the `r order_1$order_hour_of_day`th hour of the day (10am), and `r order_1$days_since_prior_order` days since the user last placed an Instacart order.  

### Aisles

```{r}
aisle_number = instacart %>% 
  distinct(aisle) %>% 
  nrow()
```
There are `r aisle_number` aisles.

```{r}
most_items_aisle = instacart %>% 
  count(aisle) %>% 
  arrange(desc(n)) %>%
  top_n(3)

```

The `r most_items_aisle$aisle[1]` aisle, is where the most items have been ordered from. The `r most_items_aisle$aisle[2]` aisle, is where the second most items have been ordered from. The `r most_items_aisle$aisle[3]` aisle, is where the third most items have been ordered from. 

### Aisles Plot

```{r}
aisle_plot = instacart %>%
  count(aisle) %>%
  mutate(aisle = fct_reorder(aisle, desc(n))) %>%
  filter(n > 10000) %>%
  ggplot(aes(x = aisle, y = n)) + geom_point() + labs(title = "Number of Items Ordered in Each Aisle", x = "Aisle", y = "Number of Items") + theme(axis.text.x = element_text(angle = 90, size = rel(1), hjust = 1))

aisle_plot
```
Add in explanation

### Aisles Table

```{r}
three_most_popular = instacart %>%
  filter(aisle == 'baking ingredients' | aisle == 'dog food care' | aisle == 'packaged vegetables fruits') %>%
  group_by(aisle, product_name) %>%
  summarize(
    count = n()) %>%
  top_n(3, count) %>%
  arrange(aisle, desc(count), product_name)
knitr::kable(three_most_popular)
```

### Pink Lady Apples and Coffee Ice Cream Table

```{r}
days_of_week = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")

mean_hour_order = instacart %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  select(product_name, order_dow, order_hour_of_day) %>%
  group_by(product_name, order_dow) %>%
  summarize(
    mean_hour = mean(order_hour_of_day)
  ) %>%
  mutate(order_dow = days_of_week[order_dow + 1]) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
knitr::kable(mean_hour_order, 
             digit = 2, 
             col.names = c("Product Name", "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
```

# Problem 2

## Load Data
```{r}
data("brfss_smart2010")

#View(brfss_smart2010)
```

## Data Cleaning
```{r}
brfss_data = brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  mutate(response = factor(response, 
                           levels = c("Poor", "Fair", "Good", "Very Good", "Excellent")))

#View(brfss_data)
```

## 7 or more loactions
```{r}
states_2002 = brfss_data %>%
  filter(year == 2002) %>%
  group_by(locationabbr) %>%
  distinct(locationdesc) %>%
  summarize(num_loc = n()) %>%
  filter(num_loc > 6) %>%
  pull(locationabbr)

states_2010 = brfss_data %>%
  filter(year == 2010) %>%
  group_by(locationabbr) %>%
  distinct(locationdesc) %>%
  summarize(num_loc = n()) %>%
  filter(num_loc > 6) %>%
  pull(locationabbr)

```

The states: `r states_2002` were observed at 7 or more locations in 2002. The states: `r states_2010` were observed at 7 or more locations in 2010.

## Excellent Dataset & Plot

```{r}
#construct the dataset
excellent_data = brfss_data %>%
  filter(response == "Excellent") %>%
  select(year, locationabbr, data_value) %>%
  group_by(year, locationabbr) %>%
  summarise(mean_value = mean(data_value, na.rm = T))
```

```{r}
#construct the spaghetti plot
excellent_plot = excellent_data %>%
  ggplot(aes(x = year, y = mean_value, color = locationabbr)) + geom_line(alpha = .5, ) + labs(title = "Average Value Over Time Within a State", y = "Average Value Within a State", color = "States")

excellent_plot
```

Add explanation

## Two Panel Plot

```{r}
#make dataset
two_panel_data = brfss_data %>%
  drop_na(response) %>%
  filter(locationabbr == "NY") %>%
  filter(year == 2006 | year == 2010) %>%
  select(year, locationabbr, locationdesc, data_value, response) %>%
  group_by(response) 

#make plot
two_panel_plot = two_panel_data %>%
  ggplot(aes(x = data_value, fill = response)) +
  geom_density() +
  facet_grid(year~response) +
  labs(
    title = "Distribution of Data Values for Responses Among Locations in NY State in 2006 and 2010", 
    x = "Data Value")

two_panel_plot
```

Add explanation

# Problem 3

## Load, Tidy and Wrangle Data
```{r}
accel_data = read_csv("./data/accel_data.csv")
View(accel_data)
accel_data = read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_count"
  ) %>%
  mutate(
    weekday = ifelse(day == "Saturday" | day == "Sunday", "weekend", "weekday"),
    minute = as.numeric(minute)
  )
```

## Description
Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

## Table of Total Activity Per Day

```{r}
accel_data %>%
  mutate(day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  group_by(week, day) %>%
  summarize(
    total_activity = mean(activity_count)
  ) %>%
  pivot_wider(
    names_from = day,
    values_from = total_activity
    ) %>%
  knitr::kable()
```

### Trends?
Are any trends apparent?

## Plot 24 Hr Activity for Each Day

```{r}
accel_data %>%
  mutate(day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>% 
  ggplot(aes(x = minute, y = activity_count, color = day)) + 
  geom_line() +
  labs(
    title = "Activity Over the Course of the Day",
    x = "Minute",
    y = "Activity"
  )
```

### Analysis
Describe in words any patterns or conclusions you can make based on this graph.
