---
title: "HW 3"
author: "Malvika Venkataraman"
date: "10/17/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```

# Problem 1

```{r}
#load data
library(p8105.datasets)
data("instacart")

#View(instacart)
```

## Description of Data

```{r}
#understanding the dataset
str(instacart)
nrow(instacart)
ncol(instacart)
length(unique(instacart$user_id))

order_1 <- instacart[1,]

order_1$order_id
order_1$product_id
order_1$add_to_cart_order
order_1$product_name
```

The dataset `instacart` records online grocery orders from Instacart users.  It contains `r nrow(instacart)` observations from `r length(unique(instacart$user_id))` unique users. The dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns. Some key variables are as follows:

* order_id: order identifier
* product_id: product identifier
* reordered: coded as 1 if this product has been ordered by the user in the past, 0 if otherwise
* product_name: name of the product
* order_number: the order sequence number for a user (1 = first, n = nth)
* days_since_prior_order: days since the last order, but capped at 30 days (NA of order_number = 1)

Let's take a look at a single row of data. In order `r order_1$order_id`, user `r order_1$user_id` ordered `r order_1$product_name` (product id = `r order_1$product_id`). The product is from the `r order_1$department` department, and was in the `r order_1$aisle` aisle. It was the `r order_1$add_to_cart_order`st product they added to their cart. This user has placed an order `r order_1$order_number` times before, and this product has been previously ordered by them (reordered = `r order_1$reordered`). The order was placed on a Thursday (order_dow = `r order_1$order_dow`), in the `r order_1$order_hour_of_day`th hour of the day (10am), and `r order_1$days_since_prior_order` days since the user last placed an Instacart order.  



